---
title: "Analysis of F0 data"
author: "Albert Baichen Du"
date: "2022-08-13"
output:
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    code_folding: hide
    theme: cerulean
  pdf_document:
    toc: yes
---

# Library
```{r setup, include = TRUE, warning = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(
  comment = "#>"
)
library(knitr)
library(tidyverse)
library(here)
library(broom)
library(broom.mixed)
library(writexl)
library(ggpubr)
library(gghighlight)
library(lme4)
library(lmerTest)
library(fs)
library(stringr)
library(lubridate)
```

## Read all data
```{r}
all_results <- dir_ls(here("results"))
smartphone_android <- dir_ls(here("smartphone_data"), regexp = "^.*QT.*")
smartphone_iphone <- dir_ls(here("smartphone_data"), regexp = "^.*RD.*")
ADEX_all <- dir_ls(here("adex_data"))
```

## Set up functions to clean the data
```{r}
read_F0 <- function(csvfile){
  
  raw_data <- read_csv(csvfile, col_names = TRUE) %>% 
    slice(-1) %>% 
    mutate(
      usability = recode(usability, "0." = "good"),
      xds = recode(xds, "0." = "C")) %>% 
    select(-F0_100) %>% 
    mutate(participant = str_extract(file_name, "\\d+")) %>% 
    relocate(participant, .before = file_name)
}

read_android <- function(csvfile){
  raw_data <- read_csv(csvfile, col_names = TRUE) %>% 
    mutate(participant = csvfile) %>% 
    mutate(participant = 
    str_extract_all(participant, "\\d+")) %>%
    mutate(system = "android")
}

read_iphone <- function(csvfile){
  raw_data <- read_csv(csvfile, col_names = TRUE) %>% 
    mutate(participant = csvfile) %>% 
    mutate(participant = 
             str_extract_all(participant, "\\d+")) %>%
    mutate(system = "iphone") %>% 
    mutate(start_time = as.character(start_time),
           end_time = as.character(end_time))
}

read_adex <- function(csvfile){
  raw_data <- read_csv(csvfile, col_names = TRUE) %>% 
    mutate(initial_time = Clock_Time_TZAdj[1]) %>% 
    mutate(participant = File_Name) %>% 
    separate(participant, into = c('participant', 'suffix'), sep = "_") %>% 
    select(-suffix) %>% 
    select(initial_time, participant) %>% 
    unique()  %>% 
    mutate(initial_time = mdy_hms(initial_time),
      year = year(initial_time),
      month = month(initial_time),
      day = day(initial_time)) %>% 
    mutate(initial_time = as.character(initial_time),
           participant = as.numeric(participant))
}

```

## Execute the functions with respective data
```{r, warning = FALSE, message=FALSE}
F0 <- purrr::map_dfr(all_results, read_F0)

smartphone_android_data <- purrr::map_dfr(smartphone_android, read_android)

smartphone_iphone_data <- purrr::map_dfr(smartphone_iphone, read_iphone)

smartphone_all <- rbind(smartphone_android_data, smartphone_iphone_data) %>% 
  mutate(participant = as.numeric(participant))

ADEX_all_data <- purrr::map_dfr(ADEX_all, read_adex)
```

## Add timestamp for F0 data
```{r}
F0_with_time <- F0 %>% 
  mutate(participant = as.numeric(participant)) %>% 
  left_join(ADEX_all_data, by = 'participant') %>% 
  relocate(initial_time, .before = speaker) %>% 
  mutate(initial_time = ymd_hms(initial_time)) %>% 
  mutate(interval_start_time = initial_time + interval_start,
         interval_end_time = initial_time + interval_end) %>% 
  relocate(c("interval_start_time", "interval_end_time"), .before = speaker) %>% 
  mutate(initial_time = as.character(initial_time)) 
```

## Add timestamp for smartphone usage data
```{r}
smartphone_usage <- ADEX_all_data %>% 
  right_join(smartphone_all, by = "participant")

smartphone_usage_android <- smartphone_usage %>% 
  filter(system == "android") %>% 
  mutate(start_time = paste0(year, "-", start_time),
         end_time = paste0(year, "-", end_time)) %>% 
  mutate(tz = case_when(
    year == 2020 & month <= 10 ~ "Etc/GMT-5",
    year == 2020 & month > 10 ~ "Etc/GMT-6",
    year == 2021 & month <= 2 ~ "Etc/GMT-6",
    year == 2021 & month == 3 & day < 14 ~ "Etc/GMT-6",
    year == 2021 & month == 3 & day >= 14 ~ "Etc/GMT-5",
    year == 2021 & month >= 4 & month <= 10  ~ "Etc/GMT-5",
    year == 2021 & month == 11 & day <= 6 ~ "Etc/GMT-5",
    year == 2021 & month == 11 & day >= 7 ~ "Etc/GMT-6")) 

smartphone_usage_android_GMT5 <- smartphone_usage_android %>% 
  filter(tz == "Etc/GMT-5") %>% 
  mutate(start_time = parse_date_time(start_time, '%Y-%d-%m %H:%M:%p', tz = "Etc/GMT-5"),
         end_time = parse_date_time(end_time, '%Y-%d-%m %H:%M:%p', tz = "Etc/GMT-5"))

smartphone_usage_android_GMT6 <- smartphone_usage_android %>% 
  filter(tz == "Etc/GMT-6") %>% 
  mutate(start_time = parse_date_time(start_time, '%Y-%d-%m %H:%M:%p', tz = "Etc/GMT-6"),
         end_time = parse_date_time(end_time, '%Y-%d-%m %H:%M:%p', tz = "Etc/GMT-6"))

smartphone_usage_android <- smartphone_usage_android_GMT5 %>% 
  rbind(smartphone_usage_android_GMT6) %>% 
  select(-tz)

smartphone_usage_iphone <- smartphone_usage %>% 
  filter(system == "iphone")

smartphone_usage_all <- smartphone_usage_android %>% 
  rbind(smartphone_usage_iphone)
```

## Remove some intermediate data
```{r}
rm(smartphone_usage_android)
rm(smartphone_usage_iphone)
rm(smartphone_usage_android_GMT5)
rm(smartphone_usage_android_GMT6)
rm(smartphone_usage)
rm(smartphone_android_data)
rm(smartphone_iphone_data)
rm(smartphone_all)
```

## Merge F0 and smartphone usage data
```{r}
meta <- F0_with_time %>% 
  left_join(smartphone_usage_all, by = c('participant', 'initial_time', 'year', 'month', 'day')) %>% 
  mutate(interval_start_time = as.character(interval_start_time),
         interval_end_time = as.character(interval_end_time),
         start_time = as.character(start_time),
          end_time = as.character(end_time)) %>% 
  mutate(
    match = ifelse(interval_start_time >= start_time & interval_end_time <= end_time, "match", "unmatch")
  ) %>% 
  relocate(c("year", "month", "day", "system", "match"), .before = speaker)
```

## Preliminary processing of matching between F0 and phone use
```{r}
meta_unmatch <- meta %>% 
  filter(match == "unmatch") %>% 
  select(-start_time, -end_time) %>% 
  mutate(duration = 0) %>% 
  unique() %>% 
  mutate(usage = "not_using")

meta_match <- meta %>% 
  filter(match == "match") %>% 
  select(-start_time, -end_time) %>% 
  unique() %>% 
  mutate(usage = "using")

meta <- meta_match %>% 
  rbind(meta_unmatch) %>% 
  mutate(usage_percent = duration/300) %>% 
  filter(!pitch_mean == "--undefined-- Hz") %>% 
  mutate(pitch_mean = as.numeric(str_remove_all(pitch_mean, " Hz")),
         pitch_min = as.numeric(str_remove_all(pitch_min, " Hz")),
         pitch_max = as.numeric(str_remove_all(pitch_max, " Hz")),
         pitch_sd = as.numeric(str_remove_all(pitch_sd, " Hz")))

```

## Descriptive Statistics and plots
### Preparation of summary statistics
```{r}
F0_summary <- meta %>% 
  group_by(usage) %>% 
  summarise(mean = mean(pitch_mean),
            min = min(pitch_min),
            max = max(pitch_max),
            n = length(pitch_mean),
            sd = sd(pitch_mean),
            se_pitch_mean = sd / sqrt(n),
            t_pitch_mean = qt((0.95)/2 + .5, n - 1), 
            cimin_pitch_mean = mean - t_pitch_mean * se_pitch_mean,
            cimax_pitch_mean = mean + t_pitch_mean * se_pitch_mean)

F0_summary_by_participant <- meta %>% 
  group_by(usage, participant) %>% 
  summarise(mean = mean(pitch_mean),
            min = min(pitch_min),
            max = max(pitch_max),
            n = length(pitch_mean),
            sd = sd(pitch_mean),
            se_pitch_mean = sd / sqrt(n),
            t_pitch_mean = qt((0.95)/2 + .5, n - 1), 
            cimin_pitch_mean = mean - t_pitch_mean * se_pitch_mean,
            cimax_pitch_mean = mean + t_pitch_mean * se_pitch_mean)

```

### Barplot with error bar
```{r}
F0_bar <- F0_summary %>% 
  ggplot()+
  geom_col(aes(x = usage, y = mean, fill = usage, color = usage), width = 0.4)+
  geom_errorbar(aes(x = usage, ymin = cimin_pitch_mean, ymax = cimax_pitch_mean), width = 0.2)+
  theme_pubclean()+
  labs(
    title = "Global Comparison of F0 When Using and not Using Cell Phones"
  )+
  theme(aspect.ratio = 3/4,
        legend.position = "right",
        plot.title = element_text(size = 13,
                                  face = "bold",
                                  hjust = 0.5), # adjust the plot title
        plot.subtitle = element_text(size = 15,
                                  hjust = 0.5),
        axis.title.x = element_text(size = 15,
                                    hjust = 0.5), # adjust the x-axis title 
        axis.title.y = element_text(size = 15,
                                    vjust = 0.5), # adjust the y-axis title
        axis.text.x = element_text(size = 12,
                                    hjust = 0.5), # adjust the x-axis labels
        axis.text.y = element_text(size = 10,
                                   vjust = 0.5), # adjust the  y-axis labels
        axis.ticks.x = element_blank() # removing the x-axis tick lines
  )
F0_bar
```

### Correlation between percent of usage and F0
There are so few data points now. Probably we can do this after all recordings are annotated.
```{r}
F0_scatter <- meta %>% 
  ggplot(aes(x = pitch_mean, y = usage_percent))+
  geom_point()+
  geom_smooth(method = "lm")+theme_pubclean()+
  labs(
    title = "Scatterplot of pitch mean against usage percent of smartphone"
  )

F0_scatter
```

## Inferential Statistics
### T-test
Number of samples is different. Paired T-test within participant might not be useful. So I did cross-participant paired t-test. For now, there is not enough data points to do this, but we will see it later.
```{r}
F0_summary_by_participant_using <- F0_summary_by_participant %>% 
  filter(usage == "using")

F0_summary_by_participant_notusing <- F0_summary_by_participant %>% 
  filter(usage == "not_using")

F0_ttest <- t.test(F0_summary_by_participant_using$mean, F0_summary_by_participant_notusing$mean)

```

### Mixed Effects Linear Model - Full Interaction
```{r}
lm_1 <- lmer(pitch_mean ~ xds * usage * Child_Age * Child_Gender + (1|participant), data = meta)
summary(lm_1)
```